### 2020-10-12 doc/manuscript/Dow_manuscript.Rmd
Fleshing out discussion outline based on current results

### Discussion

This analysis is one of a first generation of intraannual tree growth studies and as such, is the metaphorical tip of the ice berg. Here, we focused our analysis purely on early Spring warming. This falls short of the reality of our changing climate as every season if affected in some way. For example, diffuse porous trees seemed to show little, if any, relationship with the above climate window, but if we shift that window forward 20 days, the effect is much more severe, signalling a greater impact of May/June TMAX on trees of that wood type (figure or table?). A paper by Helcoski et al (2019) found that trees at SCBI had a significantly negative correlation between tree growth and May temperatures, agreeing with the results here. Other studies have found that Autumn leaf senescence  is being delayed, potentially in response to warmer temperatures in the previous year's Winter or Autumn (), this warming also impacts chilling requirements as well (), total growth is being stunted by drought and high TMAX in the current year's Summer (). How these realtionships may impact intraannual growth patterns or variables like thos described here is uncertain, but identifying such relationships will help us further understand how climate change is impacting our forests.

>>>>

### Discussion
-Growth timing
We have shown that intraannual growth patterns in North American temperate trees change in response to increasing Spring temperatures. The magnitude of these changes depends largely on the evolutionary differences inherent in the two wood types analyzed here. Ring-porous trees must repair winter embolism of their large xylem vessels before leaf expansion, achieving roughly 25% total yearly growth simultaneously with bud burst, and 50% yearly growth around the time of 75% leaf emergence (D'Orangeville et al,). On the other hand, diffuse-porous trees are less prone to cavitation and so do not need to repair winter embolism before bud burst, allowing up to 75% leaf out before reaching their 25% yearly growth totals (D'Orangeville et al). These differences are likely evolutionary hydraulic adaptations to climate(). The presence of wide-xylem cells in ring-porous trees allows for greater water conductivity in the Spring at the cost of increased cavitation risk during Spring drought (). The pre-leaf stem growth occurring in these species is fueled by carbon stores often used for other cell maintenance activities within the plant (). The dependence on these carbon stores for cambial activation could prove to be a detriment to the tree, as exhaustion of reserves leads to carbon starvation (). Diffuse-porous trees have narrower xylem cells lowering the risk of cavitation, but leading to a decreased water conductivity in the Spring, focusing on a faster, shorter growing season in the Summer. This allows the diffuse-porous trees to conserve their carbon stores, making them potentially more resistant to pest/pathogen invasions?, but exposing them to higher temperatures and drought risk during their Summer focused growth period.

The mechanisms controlling activation of cambium reactivation is still a mystery, as some speculate Auxin, the chemical believed to kick-start the reactivation process, originates in the leaves(). However, the fact that ring-porous species complete 25% growth before the presence of leaf buds indicates there may be another source of Auxin production.

Ring-porous trees reach 25% and 50% growth milestones earlier (insert actual # days?) in the year, while 75% remains statistically unchanged, in response to warmer temperautres during early Spring. This implies that the growing season of ring-porous trees expanded due to Spring temperatures, which are projected to continue to increase as anthropogenic global warming intensifies. The reactivation of cambial cells is likely beginning earlier potentially exposing these trees to higher risk of cavitation or leaf damage during frost events (). Damage of this nature would then need to be repaired using carbon reserves which may be in short demand as they are already being used for initial growth. Hypothetically, this could lead to carbon starvation in some trees. Meanwhile, diffuse-porous growth timing seems relatively unaffected by early Spring temperatures, instead showing a higher correlation with late Spring or Summer temps. This disconnect means that diffuse-porous trees will continue tofocus their growing season in the Summer, which is projected to become less condusive to tree growth as temperatures and drought intensity increase. However, diffuse-porous trees are well adapted to drought conditions, and may continue to be successful even in future climate scenarios. If not, and they begin to suffer without being able to adapt by shifting their growing period earlier, the range of some of these species may migrate North to cooler summer conditions becoming potentially extant in the Southern / mid Atlantic of the USA.    (Is mortality of ring-porous species higher in recent decades? Hard to pinpoint, but would likely be caused by warmer early-spring followed by late frost events exhausting NSC reserves?)

This analysis is one of a first generation of intraannual dendroband tree growth studies, and as such, is the metaphorical tip of the ice berg. Here, we focused our analysis purely on early Spring warming. This falls short of the reality of our changing climate, where each season is being affected in some way, leading to a combination of stressors on plant growth. For example, diffuse porous trees seemed to show little, if any, relationship with the climate window defined here, but if we shift that window forward 20 days, the effect increases dramatically, signalling a greater importance of May/June TMAX on trees of that wood type (figure or table?). This revelation agrees with the results from Helcoski et al (2019), which found differences in interspecific growth responses to several climate variables throughout the year using tree cores. Additionally, Other studies have found that Autumn leaf senescence is being delayed potentially in response to warmer temperatures in the previous year's Winter or Autumn (), changes in chilling requirements due to current year Autumn and winter temperatures (), and stunted total productivity induced by drought and high TMAX in the current year's Summer (). How these relationships may impact intraannual growth patterns or variables like those described here is uncertain, but identifying such relationships will help us further understand how climate change is impacting forests.
